---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(excess)
```

# excess

<!-- badges: start -->
<!-- badges: end -->

The `excess` package is designed to provide a set of tools and workflows to analyze and interpret hydrological cross-sectional data. This package is currently in development, so use with caution :smile:.

## Installation

You can install the development version of excess from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("brownhr/excess")
```

## Calculating cross-sectional area

Under the hood, the `xs_area()` function calculates the area of a cross-section through use of the **trapezoidal rule**, as provided by the `pracma` package.


```{r testdata}
library(excess)
library(units)
library(dplyr)

head(testxs)

```


```{r data, eval = F}
## If you can't find the `testdata` object, load it with data()

data(testxs)
```



```{r xs-area}

testxs %>% 
  xs_area(tape = TAPE,
          depth = InvertRod,
          baseline = Bankful,
          sum_area = T)

```

You can visualize cross-sections with `xs_ribbon()`


```{r xs-ribbon}

xs_ribbon(testxs)

```


For more information, check out the vignettes by navigating to https://brownhr.github.io/excess

