---
title: "xs-area"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xs-area}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Calculating the cross-sectional area of a stream

For this guide, we'll take two approaches for calculating the area: one for non-georeferenced data (i.e., just a table with your measurements, as seen in `testxs`), and one for georeferenced data (made by `sf`).


## Non-georeferenced data

Suppose you have a dataset like that which exists in `testxs`:

```{r setup}
library(excess)
library(dplyr)
library(ggplot2)
library(units)
```

```{r testxs}
head(testxs)
```
You'll notice that the data has linear units listen in `[ft]`. This comes from the [units](https://cran.r-project.org/web/packages/units/index.html) package. Units of measurement aren't necessary -- the functions work fine without them -- but it's definitely good practice to ensure you're calculating accurate results!

Anyways, we wish to calculate the cross-sectional area of the stream, and we have both the "x coordinate" (`TAPE`) and the "z-coordinate" (`InvertRod`). The `Bankful` column represents the bankful height, which is the "reference" for the depth measurement.

To calculate the cross-sectional area, `excess` uses the Trapezoidal Rule:


```{r trap-area, warning = FALSE, echo = FALSE}
testxs$depth_baseline <- pmin(testxs$InvertRod, testxs$Bankful) - testxs$Bankful

testxs %>% 
  ggplot(aes(x = TAPE, y = depth_baseline)) + 
  geom_path() + 
  geom_point() + 
  # geom_vline(aes(xintercept = TAPE), linetype = "dashed") + 
  geom_area(fill = "blue",
            alpha = 0.25) + 
  geom_segment(aes(x = TAPE, xend = lead(TAPE), 
                   y = pmax(depth_baseline, lead(depth_baseline)), 
                   yend = pmax(depth_baseline, lead(depth_baseline))),
               color = "blue",
               alpha = 0.5,
               linetype = "dashed",
               size = 1) + 
  geom_segment(aes(x = TAPE, xend = TAPE,
                   y = 0, yend = depth_baseline),
               color = "blue",
               alpha = 0.5,
               linetype = "dashed",
               size = 1) + 
  labs(x = "Tape",
       y = "Baseline Depth") + 
  theme_light()

```


The trapezoidal area can be defined as:

$$A = \frac{1}{2} (x_n - x_{n-1})(y_n + y_{n-1})$$

For an implementation in R, we can use the `dplyr::lag()` and `dplyr::lead()` functions. 


